An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This work has been partially funded by CoMeDiC ANR-15-CE40-0006 research grant.}

### 2019-07-08 jmiv.tex
Relecture 1 finie

Geometric quantities are particularly useful as regularizers in low-level image analysis, especially when little prior
information is known about the shape of interest. Length penalization is a \HT{well-behaved,} general purpose
regularizer and \HT{many} models \HT{in the literature} make use of it, from active contours \cite{casseles97geodesic}
to level-set formulations \cite{malladi1995image,malladi1995shape}. Discrete graph-based variational models have been
particularly succesfull to incorporate length penalization as a penalizer, while keeping the ability to extract \HT{a}
global optimum \cite{boykov01graphcut,appleton05geodesic}.

However length regularization shows limitations when segmenting small or thin and elongated objects, as it tends to
shrink solutions or \HT{yields} disconnected solutions. Such drawbacks can be problematic in image segmentation, image
restoration or image inpainting. It is thus rather natural to consider curvature (especially squared curvature) \HT{as}
a potential regularizer. Energies involving \HT{both length and} squared curvature are often called \HT{the {\em Elastica}}
model (\HT{that} dates back to Euler) . \HT{The {\em Willmore energy} is its $n$-dimensional extension. These were brought to
attention} in computer vision by Mumford \cite{mumford1994elastica}. A similar concept is also present in the
second-order regularizer of the \HT{original} snake model \cite{kass1988snakes}. \HT{Indeed, an explicit curvature
  term was often employed in early level-set methods, like~\cite{malladi1995image,malladi1995shape,ballester01filljoint}
  for segmentation or inpainting,} \HT{However, Casselles {\em et al.} in~\cite{casseles97geodesic} observed that these
  terms as employed were not geometric, in the sense that they depended on the discretization parameters}.

\subsection{Existing works}
\HT{The use of curvature for surface regularization in the case of the Willmore energy was studied
  in~\cite{bobenko2005discrete}, although results look promising, authors used an inexact linearized version of the
  square curvature at every step.}

One of the first successful uses of \HT{geometric, squared} curvature in image processing is the inpainting algorithm
described in \cite{masnou98inpainting}\HT{, where authors} evaluate
the absolute curvature along the level lines of a
simply-connected image to reconstruct its occluded parts. The non-intersection property of level lines \HT{induces} the
construction of an efficient dynamic programming algorithm. \HT{However, } the curvature energy is \HT{only} coarsely
approximated.  \cite{chan02elasticainpainting} \HT{is} another geometric inpainting method involving elastica
\HT{optimized with level-sets}.

In image segmentation, Zehiry and Grady have shown that injecting curvature as a regularizer can help recover thin
and elongated objects \cite{zehiry10fast}. Similarly Schoenemann {\em et al.}  \cite{schoenemann09linear} have
considered the elastica energy in their ratio-based image segmentation approach. In~\cite{schoenemann2011elastic},
Schoenemann~{\em et al.} extend ratio-cut approaches to segmentation with a curvature term. Their global optimization
framework shows the considerable \HT{advantages} of using such regularizer in common binary image segmentation tasks.
However the time complexity of their algorithm \HT{is prohibitive,} even if curvature is \HT{again} coarsely
approximated.  \HT{In more recent work, Lim {\em et al.} propose an edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set formulation. They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but observe that a shape prior and a close-enough
solution must also be provided to achieve good results.}

In fact, it is still a very challenging task to \HT{efficiently handle} curvature in the context of image segmentation.
\HT{State-of-art methods are in practice} difficult to optimize and \HT{do not scale easily}
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In order to achieve
reasonable running times, such approaches make use of coarse curvature estimations for which the approximation error is
\HT{often} unknown. Improving the quality of the curvature estimator has an important impact on the accuracy of the
results, but is computationally too costly in these methods.

Segmentation methods using Elastica or Willmore energies \HT{can be formulated via the computation of} corresponding
flow, i.e. following its gradient. Well-founded level-set methods involving Willmore energy \cite{droske2004level} have
been used in medical image segmentation \cite{lim2012introducing}. Another level-set formulation similar to the
Chan-Vese model \HT{was} proposed in \cite{zhu2013image}. Willmore flow can also be carried out by phase-field models
(see \cite{bretin2015phase}) but they are less suited to image segmentation since interfaces are blurred in such
models. Threshold dynamics can also be considered for such energies \cite{esedoglu2008threshold} and has been proposed
for image desocclusion \cite{esedoglu2005threshold}. \HT{However, flow-based methods will always deliver a local
  minimum, which heavily depend on initialization for good results}.  \HT{A similar approach is to formulate the
  Elastica as a regularization in a variational framework and use operator splitting to optimize it. Tai {\em et
    al.}~\cite{tai11elastica} and more recently Deng~{\em et al.}~\cite{deng2019new} have proposed this approach, which
  is well-suited to inverse problem solving. The problem remains non-convex and the nature of the converged iterate
are not studied, i.e. how close they are to a global optimum.}

\HT{In contrast, researchers have sought convex relaxation of curvature-related
  formulation. In~\cite{goldluecke11totalcurvature}, Goldluecke and Cremers introduce the {\em Total Curvature}, based
  on the Menger-Melnikov curvature of a measure. The formulation is non-convex but can be relaxed into a well-correlated
  but non-tight convex one. The relaxed formulation can be used to solve inverse problems including segmentation,
  however computing a solution is still expensive, requiring signifiant parallelisation efforts with a GPU to achieve
  acceptable computing times}. \HT{In~\cite{bredies15convex}, Bredies {\em et al.} study a convex approximation of the
  Elastica energy, using a lifting scheme. Thanks to the lifting, it is not restritcted to binary sets. It shows
  promising results and can be used for image restoration in addition to inpainting and segmentation. The scheme is
  however neither tight nor exact, and required 4 dimensions for a 2D approximation. In addition, it does not carry over
to $nD$. It also remains computationally expensive.}

\HT{Discrete approaches have been used to tackle the Elastica difficulties. As noted before~\cite{schoenemann09linear}
  have linearized a mesh-based approach, which was expanded upon by~\cite{strandmark11globalframework}, while
  \cite{zehiry10fast} have used a binary solver for a non-submodular approach on a regular square grid of pixels, which
  was expanded in~\cite{el2016contrast}. In~\cite{nieuwenhuis14efficient} , a specific solver is developed to yield a
  better solution to an extension of the formulation in~\cite{zehiry10fast}.}

\subsection{Motivation}
\HT{In spite of the multiplicity of existing approaches, it seems none is currently completely satisfactory. While the
  sought solution is expected to be very smooth based on geometrical expectations, the mathematical properties
  of the related operators make this smoothness elusive. Non-convexity and the presence of high-order terms make
  Elastica difficult to optimize and to approximate, let alone compute efficiently. The computed solution in the
  literature are rarely evaluated, and the complexity of the proposed solution is often high.}

@@

\HT{Consequently,} we are interested in \HT{studying} purely discrete variational segmentation models involving an
elastica energy. Recently, new estimators of curvature along digital contours have been proposed
\cite{roussillon11mdca,coeurjolly13integral,schindele17mdca} with the desirable multigrid convergence property. This
motivates us to propose models in which they can be used successfully.

\HT{In the remainder, we propose to} investigate the use of the digital integral invariant curvature estimator
\cite{coeurjolly13integral} in a discrete variational segmentation framework. More precisely, we show how to incorporate
it in a digital flow minimizing its squared curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO algorithm \cite{rother07qpbo}. \HT{Our solution
  does not} compute a global solution to the whole domain, but our approach leads to a digital flow similar to
continuous \HT{Elastica}, as shown by our experiments. We present an application of the model as a post-processing step
in a segmentation framework, and demonstrate how it can improve standard results. 
\daniel{This paper is an extended version of a previous work that appeared in \cite{antunes19}. We've included a deeper
  discussion on the optimization method we've used and we present a comparison of our segmentation method with two other
  related works, and illustrate the results with several examples. Moreover, we have included a new local combinatorial
  optimization approach for elastica minimization. Finnaly,  the code produced in this paper is freely available on
  github\footnote{https://www.github.com/danoan/BTools}.}

\subsection{Outline}.
Section two reviews the concept of multigrid convergence and highlights its importance for the
definition of digital estimators. Next, we describe two convergent estimators used in this paper, one for tangent and
the other for curvature. They are used in the optimization model and in the definition of the digital elastica. Section
three describes a local combinatorial optimization model suitable for different curvature estimators. Section four
describes the proposed curvature evolution model along with several illustrations of digital flows. Section five
explains how to use the evolution model as a post-processing step in an image segmentation framework.

>>>>

Geometric quantities are particularly useful as regularizers in
low-level image analysis, especially when little prior information is
known about the shape of interest. Length penalization is a
well-behaved, general purpose regularizer and many models in the
literature make use of it, from active contours
\cite{casseles97geodesic} to level-set formulations
\cite{malladi1995image,malladi1995shape}. Discrete graph-based
variational models have been particularly succesful to incorporate
length penalization as a penalizer, while keeping the ability to
extract a global optimum \cite{boykov01graphcut,appleton05geodesic}.

However length regularization shows limitations when segmenting small
or thin and elongated objects, as it tends to shrink solutions or
yields disconnected solutions. Such drawbacks can be problematic in
image segmentation, image restoration or image inpainting. It is thus
rather natural to consider curvature (especially squared curvature) as
a potential regularizer. Energies involving both length and squared
curvature are often called the {\em Elastica} model (that dates back
to Euler). The {\em Willmore energy} is its $n$-dimensional
extension. These were brought to attention in computer vision by
Mumford \cite{mumford1994elastica}. A similar concept is also present
in the second-order regularizer of the original snake model
\cite{kass1988snakes}.Indeed, an explicit curvature term was often
employed in early level-set methods,
like~\cite{malladi1995image,malladi1995shape,ballester01filljoint} for
segmentation or inpainting. However, Casselles {\em et al.}
in~\cite{casseles97geodesic} observed that these terms as employed
were not geometric, in the sense that they depend on the
discretization parameters.

\subsection{Existing works}
The use of curvature for surface regularization in the case of the
Willmore energy was studied in~\cite{bobenko2005discrete}. Although
results look promising, authors used an inexact linearized version of
the square curvature at every step.

One of the first successful uses of geometric, squared curvature in
image processing is the inpainting algorithm described in
\cite{masnou98inpainting}, where authors evaluate the absolute
curvature along the level lines of a simply-connected image to
reconstruct its occluded parts. The non-intersection property of level
lines induces the construction of an efficient dynamic programming
algorithm. The curvature energy is nevertheless only coarsely
approximated. We may quote \cite{chan02elasticainpainting} as another
geometric inpainting method involving elastica, which is optimized with
level-sets.

In image segmentation, Zehiry and Grady have shown that injecting
curvature as a regularizer can help recover thin and elongated objects
\cite{zehiry10fast}. Similarly Schoenemann {\em et al.}
\cite{schoenemann09linear} have considered the elastica energy in
their ratio-based image segmentation
approach. In~\cite{schoenemann2011elastic}, Schoenemann~{\em et al.}
extend ratio-cut approaches to segmentation with a curvature
term. Their global optimization framework shows the considerable
advantages of using such regularizer in common binary image
segmentation tasks.  However the time complexity of their algorithm
is prohibitive, even if curvature is again coarsely
approximated.  In a more recent work, Lim {\em et al.} propose an
edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set
formulation. They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but
observe that a shape prior and a close-enough initialization must also be
provided to achieve good results.

In fact, it is still a very challenging task to efficiently handle
curvature in the context of image segmentation.  State-of-art methods
are in practice difficult to optimize and do not scale easily
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
often unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods.

Segmentation methods using Elastica or Willmore energies can be
formulated via the computation of the corresponding flow,
i.e. following its gradient. Well-founded level-set methods involving
Willmore energy \cite{droske2004level} have been used in medical image
segmentation \cite{lim2012introducing}. Another level-set formulation
similar to the Chan-Vese model was proposed in
\cite{zhu2013image}. Willmore flow can also be carried out by
phase-field models (see \cite{bretin2015phase}) but they are less
suited to image segmentation since interfaces are blurred in such
models. Threshold dynamics can also be considered for such energies
\cite{esedoglu2008threshold} and has been proposed for image
desocclusion \cite{esedoglu2005threshold}. However, flow-based methods
will always deliver a local minimum, which heavily depend on
initialization for good results. A similar approach is to formulate
the Elastica as a regularization in a variational framework and use
operator splitting to optimize it. Tai {\em et
  al.}~\cite{tai11elastica} and more recently Deng~{\em et
  al.}~\cite{deng2019new} have proposed this approach, which is
well-suited to inverse problem solving. The problem remains non-convex
and the nature of the converged iterates are not studied, i.e. how
close they are to a global optimum.

In contrast, researchers have sought convex relaxation of curvature-related
formulation. In~\cite{goldluecke11totalcurvature}, Goldluecke and Cremers introduce the {\em Total Curvature}, based
on the Menger-Melnikov curvature of a measure. The formulation is non-convex but can be relaxed into a well-correlated
but non-tight convex one. The relaxed formulation can be used to solve inverse problems including segmentation. However computing a solution is still expensive, requiring signifiant parallelisation efforts with a GPU to achieve
  acceptable computing times. In~\cite{bredies15convex}, Bredies {\em et al.} study a convex approximation of the
  Elastica energy, using a lifting scheme. Thanks to the lifting, it is not restricted to binary sets. It shows
  promising results and can be used for image restoration in addition to inpainting and segmentation. The scheme is
  however neither tight nor exact, and required 4 dimensions for a 2D approximation. In addition, it does not carry over
to $nD$. It also remains computationally expensive.

Discrete approaches have been used to tackle the Elastica
difficulties. As noted before, \cite{schoenemann09linear} have
linearized a mesh-based approach, which was expanded upon
by~\cite{strandmark11globalframework}, while \cite{zehiry10fast} have
used a binary solver for a non-submodular approach on a regular square
grid of pixels, which was expanded
in~\cite{el2016contrast}. In~\cite{nieuwenhuis14efficient} , a
specific solver is developed to yield a better solution to an
extension of the formulation in~\cite{zehiry10fast}.

\subsection{Motivation}
In spite of the multiplicity of existing approaches, it seems none is
currently completely satisfactory. While the sought solution is
expected to be very smooth based on geometrical expectations, the
mathematical properties of the related operators make this smoothness
elusive. Non-convexity and the presence of high-order terms make
Elastica difficult to optimize and to approximate, let alone compute
efficiently. The computed solutions in the literature are rarely
evaluated, and the complexity of the proposed solution is often high.

@@

Consequently, we are interested in studying purely discrete
variational segmentation models involving an Elastica
energy. Recently, new estimators of curvature along digital contours
have been proposed
\cite{roussillon11mdca,coeurjolly13integral,schindele17mdca} with the
desirable multigrid convergence property. This motivates us to propose
models in which they can be used successfully.

In this paper, we propose to investigate the use of the digital
integral invariant curvature estimator \cite{coeurjolly13integral} in
a discrete variational segmentation framework. More precisely, we show
how to incorporate it in a digital flow minimizing its squared
curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO
algorithm \cite{rother07qpbo}. Our solution does not compute a global
solution to the whole domain but a global solution to a band
surrounding a given shape. Our approach thus leads to a digital flow
similar to continuous Elastica, as shown by our experiments. We
present an application of this model as a post-processing step in a
segmentation framework, and demonstrate how it can improve standard
results. This paper is an extended version of a previous work that
appeared in \cite{antunes19}. We have included a deeper discussion on
our optimization method and we present a more thorough comparison of
our segmentation method with two other related works. Moreover, we
have included a new local combinatorial optimization approach for
elastica minimization, which show that our discrete model is
well-founded. Finally, the code produced in this paper is freely
available on github.\footnote{https://www.github.com/danoan/BTools}

\subsection{Outline}
Section~2 reviews the concept of multigrid convergence and highlights
its importance for the definition of digital estimators. We describe
two convergent estimators used in this paper, one that approaches the
length of elementary discrete contour elements, and the other that
approaches the curvature of a discrete contour. They are used in the
optimization model and in the definition of the digital
elastica. Section~3 describes a local combinatorial optimization model
suitable for different curvature estimators, that presents interesting
results but is too costly in practice. Section~4 describes the
proposed Elastica-driven curvature evolution model along with several
illustrations of digital flows. Section~5 explains how to use this
evolution model as a post-processing step in an image segmentation
framework and compares it to two related approaches. We conclude and
draw some perspectives to this work in Section~6.
